\section{Abstract}\label{sec: abstract}
This thesis accesses the representation of configuration knowledge for customizable products in Answer Set Programming (ASP). The thesis consists of two parts. 
First, a systematic literature review is conducted to identify and analyze the main principles of product configuration.
The results of the literature review are used to create a configurator that can represent configuration knowledge independent from the domain knowledge using the declarative programming language ASP. 
The domain or the product knowledge is represented in a fact-format, in which the product experts can represent the knowledge without being ASP experts. 
This format also allows the mapping of product knowledge from other systems. 
The configuration knowledge represents general constraints that were identified in the literature survey like the incompatibility constraints, structural requirements, user requirements among others.
For a valid configuration solution, all the constraints have to be satisfied.
A bike configuration problem is used to illustrate and test the encoding. 
Therefore, three use cases are introduced to explain how user requirements are included in the encoding and how the configured product is represented. 
For future work, compatibility constraints, resource constraints and default values could be implemented. 
For an optimal solution, ASP optimization criteria such as minimize or maximize could be applied.
Methods for interactive configuration and configuration diagnosis could also be offered.  


